---
title: "$\beta$ Diversity Plots"
output:
  pdf_document: default
  html_notebook: default
---

# Pure $\beta$ Diversity

```{r}
require(tidyr)
require(dplyr)
require(ggplot2)

setwd("~/projects/QB-2021-project/")
```

# Sørensen Distance (species level)

### Heatmap of resemblance

```{r}
library(viridis)
# load the presence absence species level data
lbs <- readr::read_csv("~/projects/QB-2021-project/data/layer_by_species.csv")
layers <-  lbs[,1]
pres_abs <- lbs[,2:ncol(lbs)]
print(pres_abs)

lbs_sor <- vegan::vegdist(pres_abs, method = "bray", binary = TRUE)
print(11:1)
lattice::levelplot(as.matrix(lbs_sor)[,11:1], aspect = "iso", 
                   col.regions = viridis, xlab = "Layer", ylab = "Layer",
                   main = "Sørensen")
```
### Time series

The Other thing we can do with the presence absence matrix is produce that time series thing of species richness with the trend and jaccard distances included as well.

```{r}
# modified from https://jootse84.github.io/notes/jaccard-index-calculation-in-R#:~:text=The%20Jaccard%20Index%20is%20a,the%20length%20of%20its%20union

calculate_jaccard_rows <- function(df, row1, row2) {
  pa_mat <- df %>%
    select(-age_ma, -layer) %>%
    as.matrix()
  
  sums <- colSums(pa_mat[c(row1, row2),])
  
  inters <- length(sums[sums == 2])
  row1_s <- sum(pa_mat[row1,])
  row2_s <- sum(pa_mat[row2,])
  uni <- length(sums[sums > 0])
  
  # uni <- row1_s + row2_s - inters
  
  return(inters/uni)
}

j_vals <- c(NULL)
for (i in 1:(nrow(lbs) - 1)) {
  j_vals <- c(j_vals, calculate_jaccard_rows(lbs, i, i+1))
}

rich_series <- lbs %>%
  mutate(richness = rowSums(select(., -layer, -age_ma))) %>%
  mutate(age = age_ma) %>%
  mutate(measure = "richness") %>%
  mutate(value = richness) %>%
  select(age, measure, value)

jaccard_df <- lbs %>%
  mutate(age = age_ma) %>%
  mutate(value = c(0, j_vals)) %>%
  mutate(measure = "jaccard") %>%
  select(age, measure, value)
  
rich_jaccard <- bind_rows(list(rich_series, jaccard_df))
rich_jaccard <- spread(rich_jaccard, measure, value)

g <- ggplot(rich_jaccard, aes(x = age, y = richness, color = jaccard)) +
  stat_smooth(formula = 'y ~ x', method = 'lm', linetype='dashed', color='black') +
  geom_point(size=5) +
  scale_x_reverse() +
  xlab('Age (Million Years Ago)') +
  ylab('S') +
  scale_color_gradient(low="black", high="red") +
  theme_bw()
g
```

So we have an overall downward trend [can i get a p-value for this?] the jaccard similarity seems to be lowest following large losses in species richness.

## Bray-Curtis (coarse-grained)

### Dendrogram of sites and species
We can also produce plots using the coarse-grained family level data. We'll start with that dense as hell dendrogram heatmap thing

```{r}
# load the data
taxa <- readr::read_csv("../data/layer_by_percent.csv")

# exclude layer at age_ma cols
taxa_mat <- as.matrix(select(taxa, -layer, -age_ma))
gplots::heatmap.2(taxa_mat, distfun = function(x) vegan::vegdist(x), 
                  hclustfun = function(x) hclust(x, method="ward.D2"),
                  col = viridis, trace = "none", density.info = "none")
print(colSums(taxa_mat))
```

eh. not really an interesting and clear signal here other than some families are dominant. Bovidae is almost always a plurality in these sites.

### Ordination

We'll start with a simple PCoA before trying to get more complicated by adding environmental data.

```{r}
# do the scaling
taxa_bc <- vegan::vegdist(taxa_mat)
taxa_pcoa <- cmdscale(taxa_bc, eig = TRUE, k = 3)

# calculate explained variances
taxa_expvar <- round(sapply(taxa_pcoa$eig, function(x) x / sum(taxa_pcoa$eig)), 3)

# convert to dataframe and add some nice labels
pcoa_df <- tibble::as_tibble(taxa_pcoa$points)
nice_labels <- c("Bed I (low)", "Bed I (mid)", "Bed I (high)", "Bed II (low)", 
                 "Bed II (mid)", "Bed II (high)", "Bed III", "Bed IV (low)", 
                 "Bed IV (mid)", "Bed IV (high)", "Masek")
plot_pcoa <- pcoa_df %>%
  mutate(layer = nice_labels) %>%
  mutate(age = taxa$age_ma)
g <- ggplot(plot_pcoa, aes(x = V1, y = V2, col = age)) +
  geom_hline(yintercept = 0, linetype = 2, color = "grey") +
  geom_vline(xintercept = 0, linetype = 2, color = "grey") +
  geom_point(size = 5) +
  geom_text(aes(x = V1, y = V2 + 0.02, label = layer), color = "black") +
  scale_color_continuous(type = "viridis") +
  xlim(-0.45, 0.3) +
  ylim(-0.2, 0.3) +
  xlab(paste("PCoA 1 (", taxa_expvar[1]*100, "%)", sep="")) +
  ylab(paste("PCoA 2 (", taxa_expvar[2]*100, "%)", sep="")) +
  ggthemes::theme_pander()

g
```

It looks like we can sort of see separation between the layers with the middle time period sort of clustering in the bottom right.

